---
title: "Practical Machine Learning Course Project"
author: "Lichen Zheng"
date: "Friday, June 19, 2015"
output: html_document
---

# Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

# Data Sources

The training data for this project are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

The test data are available here: 
<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>.

# Data Processing & Prediction
## Load libraries

```{r load_library}
# Load libraries
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
```

```{r load_seed}
# Load seed
set.seed(12345)
```

## Reading Data
```{r read_csv_files}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```

## Partitioning Training Data
The training set is parted into two data sets, 60% as traing set, 40% as tesing set:
```{r data_parted}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
dim(myTraining)
dim(myTesting)
```

## Cleaning Data
Viewing possible NZV(NearZeroVariance) variables and creating a subset without NZV variables to clean NZV variables:
```{r clean_data}
# Cleaning the data
myDataNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)

myNZVvars <- names(myTraining) %in% c("new_window", "kurtosis_roll_belt", "kurtosis_picth_belt",
"kurtosis_yaw_belt", "skewness_roll_belt", "skewness_roll_belt.1", "skewness_yaw_belt",
"max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt", "avg_roll_arm", "stddev_roll_arm",
"var_roll_arm", "avg_pitch_arm", "stddev_pitch_arm", "var_pitch_arm", "avg_yaw_arm",
"stddev_yaw_arm", "var_yaw_arm", "kurtosis_roll_arm", "kurtosis_picth_arm",
"kurtosis_yaw_arm", "skewness_roll_arm", "skewness_pitch_arm", "skewness_yaw_arm",
"max_roll_arm", "min_roll_arm", "min_pitch_arm", "amplitude_roll_arm", "amplitude_pitch_arm",
"kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", "skewness_roll_dumbbell",
"skewness_pitch_dumbbell", "skewness_yaw_dumbbell", "max_yaw_dumbbell", "min_yaw_dumbbell",
"amplitude_yaw_dumbbell", "kurtosis_roll_forearm", "kurtosis_picth_forearm", "kurtosis_yaw_forearm",
"skewness_roll_forearm", "skewness_pitch_forearm", "skewness_yaw_forearm", "max_roll_forearm",
"max_yaw_forearm", "min_roll_forearm", "min_yaw_forearm", "amplitude_roll_forearm",
"amplitude_yaw_forearm", "avg_roll_forearm", "stddev_roll_forearm", "var_roll_forearm",
"avg_pitch_forearm", "stddev_pitch_forearm", "var_pitch_forearm", "avg_yaw_forearm",
"stddev_yaw_forearm", "var_yaw_forearm")
myTraining <- myTraining[!myNZVvars]
# check the new number of observations
dim(myTraining)
```

Remove first column of dataset (ID) so that it could compatible with Machine Learning Algorithms:
```{r clean_data_rm_id}

myTraining <- myTraining[c(-1)]
```

Clean Testing data set:
```{r clean_na}
training_2 <- myTraining 
for(i in 1:length(myTraining)) { 
        if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .6 ) { 
        for(j in 1:length(training_2)) {
            if( length( grep(names(myTraining[i]), names(training_2)[j]) ) ==1)  { 
                training_2 <- training_2[ , -j] 
            }   
        } 
    }
}
# check obs#
dim(training_2)

myTraining <- training_2
rm(training_2)
```

```{r clean_2}
clean1 <- colnames(myTraining)
clean2 <- colnames(myTraining[, -58]) 
myTesting <- myTesting[clean1]
testing <- testing[clean2]

# check obs #
dim(myTesting)
dim(testing)
```

```{r clean_3}

for (i in 1:length(testing) ) {
        for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(testing)[j]) ) ==1)  {
            class(testing[j]) <- class(myTraining[i])
        }      
    }      
}

testing <- rbind(myTraining[2, -58] , testing) 
testing <- testing[-1,]
```

## Machine Learning Algorithms - Random Forests
```{r ml_randomforest}
modFitB1 <- randomForest(classe ~. , data=myTraining)
predictionsB1 <- predict(modFitB1, myTesting, type = "class")
confusionMatrix(predictionsB1, myTesting$classe)
```

## Machine Learning Algorithms - Decision Tree
```{r decision_tree}
modFitA1 <- rpart(classe ~ ., data=myTraining, method="class")
fancyRpartPlot(modFitA1)
```

```{r pred}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")

confusionMatrix(predictionsA1, myTesting$classe)
```

## Prediction Results and Generating Files
```{r pred_results}
predictionsB2 <- predict(modFitB1, testing, type = "class")
predictionsB2

# Function to generate files with predictions to submit for assignment
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionsB2)
```


